---
title: "Real Estate Analysis"
author: "sho"
date: "7/28/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r data, include = FALSE, message = FALSE}

# Read data from Github to conduct analysis

library(RCurl)
library(readr)
library(dplyr)
library(purrr)
library(ggplot2)
library(tidyr)
library(knitr)

url <- getURL("https://raw.githubusercontent.com/shohct/realestate/master/RealestateData.csv")
data <- read_csv(url)
```

### Preliminary EDA on real estate data

The data used for this analysis was manually entered by the author obtained from https://www.zealty.ca/.
```{r preview, echo = FALSE}
# Preview data
kable(head(data, 5))

sqftdiff <- round(colMeans(subset(data, Bed == 2, select = Sqft)) - colMeans(subset(data, Bed == 1, select = Sqft)), 2)
```

Summary statistics on sold price, asking price, price per sqft, sqft, age, and # of days on market were examined and compared between one-bedroom apartments and two-bedroom apartments. 

The plot below shows that while both the asking and sold price of 2-bedroom condos are higher than 1-bedroom condos, when divided by sqft, the price per sqft is lower for 2-bedroom condos. This is mostly contributed by 2-bedroom condos having an overall higher square footage than 1-bedrooms (on average, 2-bedroom condos are `r sqftdiff` sqft larger than 1-bedroom condos.)

```{r EDA, echo = FALSE, fig.height = 6, fig.width = 10}
# Text review
kable(data %>% 
  select(SoldPrice, AskingPrice, Bed, Sqft, Age, DaysOnMarket, PriceSqft) %>% 
  split(.$Bed) %>% 
  map(summary))

# Plot it - need to pivot data from wide to long to use facet_wrap for all y values
data %>% 
  mutate(Bed = factor(Bed, labels = c("1-Bed", "2-Bed"))) %>% 
  pivot_longer(cols = c(SoldPrice, AskingPrice, Sqft, Age, DaysOnMarket, PriceSqft),
               names_to = "DataType",
               values_to = "Value") %>% 
  ggplot(aes(Bed, Value)) +
  geom_boxplot(aes(fill = Bed)) +
  facet_wrap(~DataType, scales = "free_y") +
  scale_y_continuous(labels = scales::comma) +
  ggtitle("Comparison of multiple variables between 1-Bed and 2-Bed apartments") +
  ylab("") +
  xlab("") +
  labs(fill = "# of Bedrooms") +
  theme(legend.position = "top",
        legend.justification = 'left')
```


